Draft -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- Draft

### 2024-10-01 docs/_tex/index.tex
add updates

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\usepackage{bookmark}

@@

Coastal areas are vital hotspots for marine biodiversity, with
intertidal seagrass meadows playing a crucial role at the interface
between the land and oceans \citep{unsworth2022}. Seagrass meadows
provide a myriad of ecosystem services, including carbon sequestration,
oxygen production, protection against sea-level rise and coastline
erosion, and mitigation of eutrophication \citetext{\citealp[
]{unsworth2022}; \citealp{sousa2019blue}}. They serve as vital habitats
for a diverse array of marine and terrestrial species, providing living,
breeding, and feeding grounds \citetext{\citealp[
]{gardner2018}; \citealp[ ]{Zoffoli2022}; \citealp{jankowska2019}}. Due
to the concentration of human activities in coastal zones, seagrass
meadows are directly exposed to and impacted by anthropogenic pressures.
Global regression and fragmentation of seagrass meadows are currently
observed due to climate change, diseases, urbanization, land
reclamation, dredging, competition with alien species, and reduction in
water quality \citetext{\citealp[ ]{nguyen2021}; \citealp[
]{soissons2018}; \citealp[ ]{orth2006}; \citealp[ ]{lin2018}; \citealp[
]{duffy2019}; \citealp[ ]{rasheed2011long}; \citealp[
]{chefaoui2018dramatic}; \citealp{sousa2019blue}}. Both habitat
fragmentation and reduction, in turn, can severely compromise the
effectiveness of ecosystem services provided by seagrass meadows. While
improvements in water quality and hydrodynamics have been recently
reported in Europe, allowing an overall recovery of seagrass ecosystems
at local and European scales, many coastal waters worldwide are still
subjected to strong eutrophication processes \citetext{\citealp[
]{deSantos2019}; \citealp[ ]{Zoffoli2021}; \citealp{sousa2019blue}}.
Coastal eutrophication has been associated to excessive accumulation of
green macroalgae, so-called green tides \citep{devlin2023nutrients}.
Green tides produce shade and suffocation over seagrass individuals,
thus threatening the health of seagrass ecosystems \citep{wang2022}.

The importance of seagrass meadows and the variety of ecosystem services
they provide have led to the enhancement of both global and regional
programs to monitor Essential Oceanic Variable (EOVs) such as seagrass
composition \citep{Miloslavich2018}, as well as Essential Biodiversity
Variable (EBVs) such as seagrass taxonomic diversity, species
distribution, population abundance, and phenology \citep{Pereira2013}.
Traditionally, indicators of seagrass status have been quantified using
\emph{in situ} measurements. However, the acquisition of these
measurements in intertidal zones is notoriously challenging. Intertidal
seagrass meadows are only exposed during low tide and can be situated in
difficult-to-reach mudflats, potentially leading to inaccurate and
limited estimations with conventional sampling techniques
\citep{nijland2019}. Satellite observations have been proven effective
in complementing \emph{in situ} sampling, allowing for the near
real-time and consistent retrieval of seagrass EOVs and EBVs over
extensive meadows. \citetext{\citealp[ ]{Zoffoli2021}; \citealp[
]{xu2021}; \citealp[ ]{Traganos2018}; \citealp{coffer2023}}

While satellite remote sensing (RS) provides temporally consistent
observations over large spatial scales, its utilization over intertidal
areas is limited by several constraints. Satellite missions with a high
temporal resolution (e.g.~daily MODIS observation) are limited by a too
coarse spatial resolution (\textgreater100 m) to accurately map patchy
seagrass meadows. Missions with a high spatial resolution such as
Sentinel-2 (10 m) or Landsat8/9 (30 m) can be limited by low spectral
resolution. The limited number of spectral bands challenges accurate
discrimination of seagrass from other co-existing macrophytes. In
particular Chlorophyceae (green algae) and marine Magnoliopsida
(seagrass) share the same pigment composition \citetext{\citealp[
]{ralph2002}; \citealp{Douay2022}}. Therefore, to someone not
specialized in the field, their spectral signatures may appear to be
alike \citetext{\citealp[ ]{Davies2023}; \citealp{bannari2022}}.
Recently, using advanced machine-learning algorithms trained with a
large hyperspectral library of more than 300 field reflectance spectra,
\citep{Davies2023} demonstrated that it was possible to discriminate
Magnoliopsida from Chlorophyceae using radiometric data acquired at
Sentinel-2 's spectral resolution. However the application of this
approach to satellite RS remains to be validated. Moreover patches of
green algae can develop at small spatial scales that are not observable
using non-commercial satellite imagery \citep{tuya2013}, especially
during the initial stage of a green tide.

Drones (Unmanned Aerial Vehicles -- UAVs) can potentially fill the data
gaps left by satellite RS and \emph{in situ} measurements, due to their
ability to provide spatially-explicit observations at very high spatial
resolutions (pixel size from mm to cm) while capturing data at
multi-spectral resolution \citetext{\citealp[
]{fairley2022drone}; \citealp{oh2017use}}. The versatility of drones
allows for their application across a diverse thematic range , from
coastal zone management \citetext{\citealp[ ]{adade2021}; \citealp[
]{casella2020}; \citealp{angnuureng2022}} to mapping species
distribution \citetext{\citealp[ ]{joyce2023}; \citealp[
]{tallam2023}; \citealp[ ]{Roca2022}; \citealp[ ]{Roman2021}; \citealp[
]{Brunier2022Topographic}; \citealp{sousa2019blue}}. However, when
applied to coastal habitat mapping, most case studies were previously
limited to a single flight, restricting the generalizability of their
application over wider geographical scales \citetext{\citealp[
]{Roman2021}; \citealp[ ]{collin2019improving}; \citealp[
]{rossiter2020uav}; \citealp{Brunier2022Topographic}}. The present study
aimed at analyzing the potential of multispectral drone RS to map
intertidal macrophytes over a diverse biogeographical range, with a
particular focus on discriminating Magnoliopsida and Chlorophyceae
(Seagrass and Green Algae, respectively). Ten drone flights were
performed over soft-bottom intertidal areas along the Atlantic
coastlines of two European countries (France and Portugal), covering a
wide range of habitats, from monospecific seagrass meadows to meadows
mixed with green, or red macroalgae. A deep learning algorithm was
trained and validated for macrophyte discrimination, emphasizing
applicability across diverse sites without a loss of prediction
accuracy. In addition, the results obtained from drone-based
multispectral remote sensing will be spatially degraded to satellite
resolutions, and the effect of this resolution on the model's
predictions will be investigated, providing insights for coastal habitat
mapping using satellites.

@@

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Neural Network model. The training pixels were
categorized into seven different classes, representing the various
habitats encountered at the different study sites: Sediment, Water,
Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and
Rhodophyceae. Only low-altitude flights (Table~\ref{tbl-flights}) were
used for training because their 8 mm spatial resolution allowed to avoid
spectral sub-pixel mixing and to accurately identify vegetation classes.
More than 418,000 pixels at 8 mm resolution from the 3 training flights
were used to train the model (Table~\ref{tbl-validationPX}). Twenty one
variables were used by the model as predictors: the ten raw spectral
bands of the Micasense RedEdge Dual MX multispectral camera (ranging
from 444 nm to 840 nm), the same ten spectral bands standardized using a
min/max transformation (Equation~\ref{eq-std} ; \citep{Cao2017}) and the
Normalized difference vegetation index (NDVI, Equation~\ref{eq-ndvi}).
Standardisation of spectral bands is commonly used to eliminate the
scaling differences between spectra and to limit the effect of biomass
on the spectra shape \citetext{\citealp[
]{Douay2022}; \citealp{Davies2023}}.

@@

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Low_res/Maps Pred/FigX-Gaf_High_Pred.png}

@@

Figure~\ref{fig-pixelsize} demonstrated the effect of pixel size on the
agreement between two products. The highest resolution at which
resampling was performed was 10 cm. At this resolution, agreement rates
had already decreased, ranging from 78\% for red algae to 94\% for
seagrass, depending on the class. As pixel size increased, the agreement
between the products steadily declined across all classes. This decline
was most pronounced for green algae, which experienced the sharpest drop
in agreement as pixel size surpassed 10 m, falling below 0\% at a 30 m
pixel size. Seagrass maintained higher agreement levels at coarser
resolutions, though a clear decline was still observed, dropping from
over 94\% at 10 cm to around 88\% at 30 m. Brown algae and red algae
showed a more moderate decline in comparison, with agreement levels at
30 m resolution settling around 62\% and 40\%, respectively.

@@

The Generalized Linear Model (GLM) analysis revealed a significant
effect of pixel resolution on the agreement, with a negative
relationship (\(\beta = -1.0334 \times 10^{-8}, p < 0.001\)). This
indicates that as pixel resolution increases, the proportion of the
response variable decreases.

The vegetation classes also exhibited distinct effects on the agreement.
Magnoliopsida showed a significant positive effect
(\(\beta = 0.1202, p < 0.001\)), suggesting that this class was
associated with a higher agreement.

@@

The primary objective of this study was to develop a method for the
accurate classification of macrophytes at low tide on tidal flats,
specifically focusing on distinguishing between Chlorophyceae (green
macroalgae) and marine Magnoliopsida (seagrasses) using a multispectral
resolution. The discrimination of seagrasses from green macroalgae
presents is spectrally challenging \citetext{\citealp[
]{oiry2021using}; \citealp[
]{bannari2022}; \citealp{veettil2020opportunities}}. These two
macrophytes share a similar pigment composition: chlorophyll-a (common
to all vegetation types), chlorophyll-b (an additional photosynthetic
pigment), and accessory carotenoids such as zeaxanthin, lutein and
neoxanthin (Figure~\ref{fig-Pigm}). Their spectral responses could be
close, particularly at a multispectral resolution. Seagrass and green
macroalgae frequently co-occur in intertidal areas, and can intermingle
within a remote sensing pixel if the spatial resolution is too low.
Here, the issue of intra-pixel mixing was resolved thanks to the very
high spatial resolution of the drone. In this study the risk of spectral
confusion was avoided with a machine-learning approach exploiting a
neural networks classifier. Our drone flights and a recent study based
on \emph{in situ} radiometry, suggested that a sensor with at least
eight spectral bands ranging from 500 to 850 nm and including a green
band at 530 nm and a RedEdge band at 730 nm was crucial to accurately
discriminate green macroalgae from seagrasses \citep{Davies2023}.

@@

Meeting these two criteria, the Micasense RedEdge-MX DUAL camera used in
this study, enabled the classifier to achieve 97\% of accuracy between
these two classes (Figure~\ref{fig-ValidationGreen}). Even if their
pigment composition is similar, differences in the spectral shape can be
observed, with green algae having a higher reflectance peak at 560 nm as
well as a higher NIR plateau than seagrass
(Figure~\ref{fig-vegetation}). Such differences were previously
attributed to differences in pigments concentration and/or ratios
\citep{bargain2013seasonal}, cellular structure as well as in the
orientation of the plant at the sediment surface \citetext{\citealp[
]{beach1997vivo}; \citealp[
]{kirk1994light}; \citealp{hedley2018influence}}.

@@

The ability to differentiate between various types of vegetation plays a
critical role in ecological monitoring and coastal management
\citep{WFD2000}. By distinguishing between seagrasses and macroalgae,
our approach facilitates targeted conservation strategies, enabling more
effective preservation and restoration efforts in coastal ecosystems.
While comparing the reflectance at two different altitudes (12 m and 120
m), a nearly one-to-one relationship was observed, with a Root Mean
Square Error (RMSE) of 0.02 (Figure~\ref{fig-CompareRef}). This result
indicates that the reflectance measured by remote sensing (RS) sensors
is not significantly influenced by pixel size. This finding is valuable
for integrating drone-based data into larger-scale mapping projects
(e.g., combining satellite and drone mapping in side-by-side analyses).
The consistency of reflectance across altitudes suggests that drones can
be effectively used for finer-scale mapping without compromising data
accuracy when merging with other platforms. However, it was observed
that there is an underestimation of the infrared part of the spectra in
the high-altitude dataset (Figure~\ref{fig-CompareRef}). Such disparity
in infrared reflectance may stem from temporal differences between the
flights, possibly resulting in a slightly drier intertidal area and
consequently higher infrared reflectance. This disparity poses an issue
for the methodology followed in the present study, relying solely on one
flight height for training. To address this issue, we employed min/max
standardized reflectance spectra as predictors for the model
Equation~\ref{eq-std}. This approach allowed us to eliminate the slight
reflectance difference between the flights (Figure~\ref{fig-CompareRef}
B) and to focus on the shape of the spectra in the visible part of the
electromagnetic spectra, where different pigmentation are associated to
taxonomic diagnostic features. In contrast to subtidal seagrasses, which
maintain a relatively constant biomass throughout the year, intertidal
seagrasses, like the one studied in this work, exhibit strong seasonal
phenology \citep{davies2024sentinel}. At some sites, they completely
disappear during the winter and reach their peak above-ground biomass in
the summer and early fall. Along with these seasonal changes in biomass,
the pigment composition and ratios also vary throughout the year,
reflecting the plants' adaptations to different environmental conditions
\citetext{\citealp[ ]{bargain2013seasonal}; \citealp{legare2022remote}}.
Standardization of spectral signatures helps to mitigate the impact of
changing biomass on the spectral profile, enabling the development of a
model that can reliably predict vegetation across different geographical
locations and seasons. This approach allows for consistent
classification of vegetation despite variations in biomass and
fluctuations in light conditions, providing a robust tool for monitoring
and predicting vegetation dynamics \citetext{\citealp[
]{fyfe2003spatial}; \citealp[
]{COSTA2021107018}; \citealp{piaser2023impact}}. We found 90 \% seagrass
cover was necessary at an 8 cm resolution for confident prediction of
its presence (Figure~\ref{fig-upscaling}). However, due to the strong
phenology of intertidal seagrass meadows in Europe, the period when a
meadow is well-established can be temporally restricted, limiting the
ideal window for accurate detection.

@@

Climate change, global warming, eutrophication, alien and invasive
species development, coastal erosion, and sea level rise are expected to
continue impacting coastal ecosystems in the future
\citetext{\citealp{SCHIBALSKI2022101414}; \citealp[
]{holon2018predictive}; \citealp{marquet2024global}} and the demand for
meaningful and efficient monitoring of coastal habitats has never been
higher\citetext{\citealp[ ]{muller2018satellite}; \citealp[
]{villalobos2023remote}; \citealp{oiry2021using}}. Our findings,
particularly the improved discrimination of intertidal seagrass and
green macroalgae from other intertidal vegetation classes, highlight the
potential of drone-based remote sensing to support diverse applications,
from conservation of biodiversity to climate change adaptation
strategies.

>>>>

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

@@

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\usepackage{bookmark}

@@

Coastal areas are vital hotspots for marine biodiversity, with
intertidal seagrass meadows playing a crucial role at the interface
between the land and oceans \citep{unsworth2022}. Seagrass meadows
provide a myriad of ecosystem services, including carbon sequestration,
oxygen production, protection against sea-level rise and coastline
erosion, and mitigation of eutrophication
\citep{unsworth2022, sousa2019blue}. They serve as vital habitats for a
diverse array of marine and terrestrial species, providing living,
breeding, and feeding grounds
\citep{gardner2018, Zoffoli2022, jankowska2019}. Due to the
concentration of human activities in coastal zones, seagrass meadows are
directly exposed to and impacted by anthropogenic pressures. Global
regression and fragmentation of seagrass meadows are currently observed
due to climate change, diseases, urbanization, land reclamation,
dredging, competition with alien species, and reduction in water quality
\citep{nguyen2021, soissons2018, orth2006, lin2018, duffy2019, rasheed2011long, chefaoui2018dramatic, sousa2019blue}.
Both habitat fragmentation and reduction, in turn, can severely
compromise the effectiveness of ecosystem services provided by seagrass
meadows. While improvements in water quality and hydrodynamics have been
recently reported in Europe, allowing an overall recovery of seagrass
ecosystems at local and European scales, many coastal waters worldwide
are still subjected to strong eutrophication processes
\citep{deSantos2019, Zoffoli2021, sousa2019blue}. Coastal eutrophication
has been associated to excessive accumulation of green macroalgae,
so-called green tides \citep{devlin2023nutrients}. Green tides produce
shade and suffocation over seagrass individuals, thus threatening the
health of seagrass ecosystems \citep{wang2022}.

The importance of seagrass meadows and the variety of ecosystem services
they provide have led to the enhancement of both global and regional
programs to monitor Essential Oceanic Variable (EOVs) such as seagrass
composition \citep{Miloslavich2018}, as well as Essential Biodiversity
Variable (EBVs) such as seagrass taxonomic diversity, species
distribution, population abundance, and phenology \citep{Pereira2013}.
Traditionally, indicators of seagrass status have been quantified using
\emph{in situ} measurements. However, the acquisition of these
measurements in intertidal zones is notoriously challenging. Intertidal
seagrass meadows are only exposed during low tide and can be situated in
difficult-to-reach mudflats, potentially leading to inaccurate and
limited estimations with conventional sampling techniques
\citep{nijland2019}. Satellite observations have been proven effective
in complementing \emph{in situ} sampling, allowing for the near
real-time and consistent retrieval of seagrass EOVs and EBVs over
extensive meadows. \citep{Zoffoli2021, xu2021, Traganos2018, coffer2023}

While satellite remote sensing (RS) provides temporally consistent
observations over large spatial scales, its utilization over intertidal
areas is limited by several constraints. Satellite missions with a high
temporal resolution (e.g.~daily MODIS observation) are limited by a too
coarse spatial resolution (\textgreater100 m) to accurately map patchy
seagrass meadows. Missions with a high spatial resolution such as
Sentinel-2 (10 m) or Landsat8/9 (30 m) can be limited by low spectral
resolution. The limited number of spectral bands challenges accurate
discrimination of seagrass from other co-existing macrophytes. In
particular Chlorophyceae (green algae) and marine Magnoliopsida
(seagrass) share the same pigment composition
\citep{ralph2002, Douay2022}. Therefore, to someone not specialized in
the field, their spectral signatures may appear to be alike
\citep{Davies2023, bannari2022}. Recently, using advanced
machine-learning algorithms trained with a large hyperspectral library
of more than 300 field reflectance spectra, \citep{Davies2023}
demonstrated that it was possible to discriminate Magnoliopsida from
Chlorophyceae using radiometric data acquired at Sentinel-2 's spectral
resolution. However the application of this approach to satellite RS
remains to be validated. Moreover patches of green algae can develop at
small spatial scales that are not observable using non-commercial
satellite imagery \citep{tuya2013}, especially during the initial stage
of a green tide.

Drones (Unmanned Aerial Vehicles -- UAVs) can potentially fill the data
gaps left by satellite RS and \emph{in situ} measurements, due to their
ability to provide spatially-explicit observations at very high spatial
resolutions (pixel size from mm to cm) while capturing data at
multi-spectral resolution \citep{fairley2022drone, oh2017use}. The
versatility of drones allows for their application across a diverse
thematic range , from coastal zone management
\citep{adade2021, casella2020, angnuureng2022} to mapping species
distribution
\citep{joyce2023, tallam2023, Roca2022, Roman2021, Brunier2022Topographic, sousa2019blue}.
However, when applied to coastal habitat mapping, most case studies were
previously limited to a single flight, restricting the generalizability
of their application over wider geographical scales
\citep{Roman2021, collin2019improving, rossiter2020uav, Brunier2022Topographic}.
The present study aimed at analyzing the potential of multispectral
drone RS to map intertidal macrophytes over a diverse biogeographical
range, with a particular focus on discriminating Magnoliopsida and
Chlorophyceae (Seagrass and Green Algae, respectively). Ten drone
flights were performed over soft-bottom intertidal areas along the
Atlantic coastlines of two European countries (France and Portugal),
covering a wide range of habitats, from monospecific seagrass meadows to
meadows mixed with green, or red macroalgae. A deep learning algorithm
was trained and validated for macrophyte discrimination, emphasizing
applicability across diverse sites without a loss of prediction
accuracy. In addition, the results obtained from drone-based
multispectral remote sensing will be spatially degraded to satellite
resolutions, and the effect of this resolution on the model's
predictions will be investigated, providing insights for coastal habitat
mapping using satellites.

@@

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Neural Network model. The training pixels were
categorized into seven different classes, representing the various
habitats encountered at the different study sites: Sediment, Water,
Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and
Rhodophyceae. Only low-altitude flights (Table~\ref{tbl-flights}) were
used for training because their 8 mm spatial resolution allowed to avoid
spectral sub-pixel mixing and to accurately identify vegetation classes.
More than 418,000 pixels at 8 mm resolution from the 3 training flights
were used to train the model (Table~\ref{tbl-validationPX}). Twenty one
variables were used by the model as predictors: the ten raw spectral
bands of the Micasense RedEdge Dual MX multispectral camera (ranging
from 444 nm to 840 nm), the same ten spectral bands standardized using a
min/max transformation (Equation~\ref{eq-std} ; \citep{Cao2017}) and the
Normalized difference vegetation index (NDVI, Equation~\ref{eq-ndvi}).
Standardisation of spectral bands is commonly used to eliminate the
scaling differences between spectra and to limit the effect of biomass
on the spectra shape \citep{Douay2022, Davies2023}.

@@

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Low_res/Maps Pred/FigX-Gaf_High_Pred1.png}

@@

Figure~\ref{fig-pixelsize} illustrated the relationship between
predicted area loss for different vegetation types and map resolution.
The highest resolution used for resampling was 10 cm, at which point
vegetation area had already started to decrease. The losses ranged from
21\% for red algae to 5\% for seagrass, depending on the vegetation
type. As pixel size increased, the retrieved vegetation area decreased
for all classes. This reduction was most severe for green algae, which
experienced a sharp decline once pixel size exceeded 10 m, culminating
in a complete loss (100\%) at a 30 m pixel size. Seagrass exhibited the
lowest loss at coarser resolutions, though it still showed a clear
decline, with area loss increasing from over 5\% at 10 cm to around 11\%
at 30 m. Brown algae and red algae experienced more moderate declines,
with losses at 30 m resolution reaching approximately 37\% and 59\%,
respectively.

@@

The Generalized Linear Model (GLM) analysis showed a significant effect
of pixel resolution on the mapped vegetation area, indicating a positive
relationship between resolution and area loss
(\(\beta = 0.067, p < 0.001\)). This suggests that as pixel resolution
increases, the mapped area of each vegetation type decreases.

@@

The primary objective of this study was to develop a method for the
accurate classification of macrophytes at low tide on tidal flats,
specifically focusing on distinguishing between Chlorophyceae (green
macroalgae) and marine Magnoliopsida (seagrasses) using a multispectral
resolution. The discrimination of seagrasses from green macroalgae
presents is spectrally challenging
\citep{oiry2021using, bannari2022, veettil2020opportunities}. These two
macrophytes share a similar pigment composition: chlorophyll-a (common
to all vegetation types), chlorophyll-b (an additional photosynthetic
pigment), and accessory carotenoids such as zeaxanthin, lutein and
neoxanthin (Figure~\ref{fig-Pigm}). Their spectral responses could be
close, particularly at a multispectral resolution. Seagrass and green
macroalgae frequently co-occur in intertidal areas, and can intermingle
within a remote sensing pixel if the spatial resolution is too low.
Here, the issue of intra-pixel mixing was resolved thanks to the very
high spatial resolution of the drone. In this study the risk of spectral
confusion was avoided with a machine-learning approach exploiting a
neural networks classifier. Our drone flights and a recent study based
on \emph{in situ} radiometry, suggested that a sensor with at least
eight spectral bands ranging from 500 to 850 nm and including a green
band at 530 nm and a RedEdge band at 730 nm was crucial to accurately
discriminate green macroalgae from seagrasses \citep{Davies2023}.

@@

Meeting these two criteria, the Micasense RedEdge-MX DUAL camera used in
this study, enabled the classifier to achieve 97\% of accuracy between
these two classes (Figure~\ref{fig-ValidationGreen}). Even if their
pigment composition is similar, differences in the spectral shape can be
observed, with green algae having a higher reflectance peak at 560 nm as
well as a higher NIR plateau than seagrass
(Figure~\ref{fig-vegetation}). Such differences were previously
attributed to differences in pigments concentration and/or ratios
\citep{bargain2013seasonal}, cellular structure as well as in the
orientation of the plant at the sediment surface
\citep{beach1997vivo, kirk1994light, hedley2018influence}.

@@

The ability to differentiate between various types of vegetation plays a
critical role in ecological monitoring and coastal management
\citep{WFD2000}. By distinguishing between seagrasses and macroalgae,
our approach facilitates targeted conservation strategies, enabling more
effective preservation and restoration efforts in coastal ecosystems.
While comparing the reflectance at two different altitudes (12 m and 120
m), a nearly one-to-one relationship was observed, with a Root Mean
Square Error (RMSE) of 0.02 (Figure~\ref{fig-CompareRef}). This result
indicates that the reflectance measured by remote sensing (RS) sensors
is not significantly influenced by pixel size. This finding is valuable
for integrating drone-based data into larger-scale mapping projects
(e.g., combining satellite and drone mapping in side-by-side analyses).
The consistency of reflectance across altitudes suggests that drones can
be effectively used for finer-scale mapping without compromising data
accuracy when merging with other platforms. However, it was observed
that there is an underestimation of the infrared part of the spectra in
the high-altitude dataset (Figure~\ref{fig-CompareRef}). Such disparity
in infrared reflectance may stem from temporal differences between the
flights, possibly resulting in a slightly drier intertidal area and
consequently higher infrared reflectance. This disparity poses an issue
for the methodology followed in the present study, relying solely on one
flight height for training. To address this issue, we employed min/max
standardized reflectance spectra as predictors for the model
Equation~\ref{eq-std}. This approach allowed us to eliminate the slight
reflectance difference between the flights (Figure~\ref{fig-CompareRef}
B) and to focus on the shape of the spectra in the visible part of the
electromagnetic spectra, where different pigmentation are associated to
taxonomic diagnostic features. In contrast to subtidal seagrasses, which
maintain a relatively constant biomass throughout the year, intertidal
seagrasses, like the one studied in this work, exhibit strong seasonal
phenology \citep{davies2024sentinel}. At some sites, they completely
disappear during the winter and reach their peak above-ground biomass in
the summer and early fall. Along with these seasonal changes in biomass,
the pigment composition and ratios also vary throughout the year,
reflecting the plants' adaptations to different environmental conditions
\citep{bargain2013seasonal, legare2022remote}. Standardization of
spectral signatures helps to mitigate the impact of changing biomass on
the spectral profile, enabling the development of a model that can
reliably predict vegetation across different geographical locations and
seasons. This approach allows for consistent classification of
vegetation despite variations in biomass and fluctuations in light
conditions, providing a robust tool for monitoring and predicting
vegetation dynamics
\citep{fyfe2003spatial, COSTA2021107018, piaser2023impact}. We found 90
\% seagrass cover was necessary at an 8 cm resolution for confident
prediction of its presence (Figure~\ref{fig-upscaling}). However, due to
the strong phenology of intertidal seagrass meadows in Europe, the
period when a meadow is well-established can be temporally restricted,
limiting the ideal window for accurate detection.

@@

Climate change, global warming, eutrophication, alien and invasive
species development, coastal erosion, and sea level rise are expected to
continue impacting coastal ecosystems in the future
\citep{SCHIBALSKI2022101414, holon2018predictive, marquet2024global} and
the demand for meaningful and efficient monitoring of coastal habitats
has never been
higher\citep{muller2018satellite, villalobos2023remote, oiry2021using}.
Our findings, particularly the improved discrimination of intertidal
seagrass and green macroalgae from other intertidal vegetation classes,
highlight the potential of drone-based remote sensing to support diverse
applications, from conservation of biodiversity to climate change
adaptation strategies.
